### 2024-12-09 manuscript.tex
Addressing comments

Global storm-resolving models (GSRMs) are the next avenue of climate modelling.
Among them is a 5-km version of the Icosahedral Nonhydrostatic Weather and
Climate Model (ICON). The high resolution means that parameterizations of
convection and clouds including subgrid-scale clouds are forgone, relying on
explicit their simulation, but still using microphysics and turbulence
parametrization.  Standard-resolution (10--100-km) models, which use
subgrid-scale convection and cloud parametrization, have substantial cloud
biases over the Southern Ocean (SO), affecting radiation and sea surface
temperature.  We evaluated SO clouds in ICON and the ERA5 and MERRA-2
reanalyses.  The SO is dominated by low clouds, which cannot be observed
accurately from space due to overlapping clouds, attenuation, and ground
clutter.  Instead, we analysed about 2400 days of lidar observations from 31
voyages and a station during 2010--2021 using a ground-based lidar simulator.
ICON and the reanalyses underestimate the total cloud fraction by about 10 and
20\%, respectively. ICON and ERA5 overestimate the cloud occurrence peak at
about 500 m, potentially explained by their lifting condensation levels being
too high.  The reanalyses strongly underestimate near-surface clouds or fog.
MERRA-2 tends to underestimate cloud occurrence at all heights.  In daily cloud
cover, ICON and the reanalyses tend to be about 1 and 2 oktas clearer,
respectively.  Conditions of weak stability are associated with the largest
errors in ICON and the reanalyses.  Compared to radiosondes, potential
temperature is accurate in the reanalyses, but ICON underestimates stability
over the low-latitude SO and too humid in the boundary layer. MERRA-2 is too
humid at all heights. SO cloud biases remain a substantial issue in the GSRM,
but are an improvement over the lower-resolution reanalyses. Explicitly
resolved convection and cloud processes (except microphysics) were not enough
to address the model cloud biases.

@@

The Southern Ocean (SO) is known to be a problematic region for climate model
biases due to a lack of surface and in situ observations and being a lower
priority region for numerical weather prediction (NWP) and climate model
development because of its distance from populated areas.  Nevertheless,
radiation biases and changes over an area of its size have a substantial
influence on the global climate, such as affecting the Earth radiation balance
and carbon cycle, and the SO is an important part of the global ocean conveyor
belt. In general, marine clouds have a disproportionate effect on top of
atmosphere (TOA) SW radiation due to the relatively low albedo of the sea
surface.  The relative longitudinal symmetry of the SO means that model cloud
biases tend to be similar across longitudes.  Here, we conventionally refer to
the SO as ocean regions south of 40°S, low-latitude SO as 40--55°S and
high-latitude SO as south of 55°S.

@@

We used the ALCF to derive cloud occurrence by height and the total cloud
fraction from observations, ICON, ERA5 and MERRA-2 (Fig.
\ref{fig:cloud-occurrence-panel}). In addition, we aggregated the data sources
(voyages and stations) by calculating the averages and percentiles of all
individual profiles, presented in Fig. \ref{fig:cloud-occurrence}. The analysis
shows that the total cloud fraction (determined as the fraction of profiles
with clouds at any height in the lidar cloud mask) is underestimated in ICON by
about 10\% and in the reanalyses by about 20\%. When analysed by height, ICON
overestimates cloud occurrence below 1 km and underestimates it above, MERRA-2
underestimates cloud occurrence at all heights, especially near the surface,
and ERA5 simulates cloud occurrence relatively well above 1 km, but strongly
underestimates it near the surface.  We note that fog or near-surface clouds
are strongly underestimated in the reanalyses (fog and clouds are both included
in the cloud occurrence).  As shown in Fig.  \ref{fig:cloud-occurrence-panel},
the biases are relatively consistent across the data sources and longitudes. We
conclude that the ICON results  the observations better than the reanalyses in
this metric.

For all observations considered (Fig. \ref{fig:cloud-occurrence}a), the data
show cloud occurrence peaking at the surface, whereas the models show a higher
peak (at about 500 m). The models generally underestimate the total cloud
fraction by 10--20\% and show a strong reduction in cloud occurrence near the
surface, which is not supported by the observations. ICON and ERA5 overestimate
cloud occurrence at their peak (between 0--1 km). Above 1 km, ICON and MERRA-2
underestimate cloud occurrence, but ERA5 is very accurate. The exaggerated peak
in models is partly supported by the lifting condensation level (LCL)
distribution, which peaks higher in the models than in the observations (at the
surface), although this is not very pronounced. This is indicative of
near-surface relatively humidity being often close to saturation in the
observations but not in the models.

@@

When subsetted by cyclonic and non-cyclonic situations (Fig.
\ref{fig:cloud-occurrence}d, e), we see that the cyclonic situations have a
larger amount of observed cloudiness, including peak and total cloud fraction.
In these situations, the model vertical profiles of cloud occurrence compare
well with observations, but still tend to underestimate cloud occurrence above
1 km and near the surface. Non-cyclonic situations are similar to the general
case, partially also because they form the majority of cases.

When subsetted by conditions of weak and strong stability (Fig.
\ref{fig:cloud-occurrence}f, g), as defined in Section
\ref{sec:cyclone-stability}, we see that in situations of strong stability
cloud occurrence peaks strongly at the surface in observations, compared to
situations of weak stability, where the peak is more diffuse between 0 and 1
km.  It is work mentioning that conditions of strong stability might be
associated with the formation of advection fog, such as in situation of warm
air advection from the north over colder sea surface, thus inducing fog
formation by cooling of the warm and humid air by the cold surface.  In
situations of strong stability, the models do a fairly good job, but
overestimate cloud occurrence at the peak below 1 km; above 1 km, they show
similar biases as in the general case.  In situations of weak stability, the
bias in ICON is very pronounced, with a much stronger peak at about 500 m, ERA5
underestimates cloud occurrence below 1 km (especially near the surface), and
MERRA-2 underestimates cloud occurrence even more strongly.

@@

In Fig.  \ref{fig:cloud-cover} we show the results for the same subsets of data
as in Section \ref{sec:cloud-occurrence}. Observations display the highest
proportion of high cloud cover values (5--8 oktas), peaking at 7 oktas. This
pattern is not represented by ICON or either reanalyses.  While ICON is closest
to matching the observed distribution, it tends to be 1 okta clearer than the
observations, peaking at 6 oktas, and substantially underestimating days with 8
oktas.  Overall, the reanalyses show results similar to each other,
underestimating cloud cover by about 2 oktas and strongly underestimating days
with 7 and 8 oktas. Of the two reanalyses, MERRA-2 shows slightly higher cloud
cover and thus is more consistent with observations.

@@

Overall, the mean $\theta_v$ is very well represented in ICON and MERRA-2,
being only slightly colder in the mid-to-high troposphere (less stable) in ICON
than in observations (Fig. \ref{fig:potential-temperature}a).  Larger
differences exist, however, in the 40--55°S zone, where ICON is colder by about
5 K at higher altitudes (Fig.  \ref{fig:potential-temperature}b). In other
subsets, the bias is relatively small. MERRA-2 is very close to the
observations, possibly due to a high accuracy of assimilation of this quantity.
Notably, the variability of virtual potential temperature (as represented by
the percentiles) is much smaller in ICON than in the observations. This
indicates that the model's internal variability in the lower-tropospheric
thermodynamic conditions in the SO is smaller than in reality.

>>>>

Global storm-resolving models (GSRMs) are the next avenue of climate modelling.
Among them is a 5-km version of the Icosahedral Nonhydrostatic Weather and
Climate Model (ICON). The high resolution means that parameterizations of
convection and clouds including subgrid-scale clouds are forgone, relying on
explicit their simulation, but still using microphysics and turbulence
parametrization.  Standard-resolution (10--100-km) models, which use
subgrid-scale convection and cloud parametrization, have substantial cloud
biases over the Southern Ocean (SO), affecting radiation and sea surface
temperature.  We evaluated SO clouds in ICON and the ERA5 and MERRA-2
reanalyses.  The SO is dominated by low clouds, which cannot be observed
accurately from space due to overlapping clouds, attenuation, and ground
clutter.  Instead, we analysed about 2400 days of lidar observations from 31
voyages and a station during 2010--2021 using a ground-based lidar simulator.
ICON and the reanalyses underestimate the total cloud fraction by about 10 and
20\%, respectively. ICON and ERA5 overestimate the cloud occurrence peak at
about 500 m, potentially explained by their lifting condensation levels being
too high.  The reanalyses strongly underestimate near-surface clouds or fog.
MERRA-2 tends to underestimate cloud occurrence at all heights.  In daily cloud
cover, ICON and the reanalyses tend to be about 1 and 2 oktas clearer,
respectively.  Conditions of weak stability are associated with the largest
errors in ICON and the reanalyses.  Compared to radiosondes, potential
temperature is accurate in the reanalyses to about 0.5 K, but ICON
underestimates stability over the low-latitude SO and too humid in the boundary
layer. MERRA-2 is too humid at all heights. SO cloud biases remain a
substantial issue in the GSRM, but are an improvement over the lower-resolution
reanalyses. Explicitly resolved convection and cloud processes (except
microphysics) were not enough to address the model cloud biases.

@@

The Southern Ocean (SO) is known to be a problematic region for climate model
biases due to a lack of surface and in situ observations and being a lower
priority region for numerical weather prediction (NWP) and climate model
development because of its distance from populated areas.  Nevertheless,
radiation biases and changes over an area of its size have a substantial
influence on the global climate, such as affecting the Earth radiation balance
and carbon cycle, and the SO is an important part of the global ocean conveyor
belt. In general, marine clouds have a disproportionate effect on top of
atmosphere (TOA) SW radiation due to the relatively low albedo of the sea
surface.  The relative longitudinal symmetry of the SO means that model cloud
biases tend to be similar across longitudes.  Here, we conventionally refer to
the SO as ocean regions south of 40°S, low-latitude SO as 40--55°S and
high-latitude SO as south of 55°S. The reason for this dividing latitude
is to split the SO into about two equal zones, as well as the results by
\cite{schuddeboom2021} (Fig. 2b), showing a contrast in CMIP model radiation
biases, \citep{schuddeboom2019} (Fig. 2) and \cite{kuma2020} (Fig. 3) showing
contrasting radiation biases in the Hadley Centre Global Environmental Model
between these two zones. The findings of \cite{niu2024}, however, support a
different dividing line of 62º based on cloud condensation nuclei
concentration.

@@

We used the ALCF to derive cloud occurrence by height and the total cloud
fraction from observations, ICON, ERA5 and MERRA-2 (Fig.
\ref{fig:cloud-occurrence-panel}). In addition, we aggregated the data sources
(voyages and stations) by calculating the averages and percentiles of all
individual profiles, presented in Fig. \ref{fig:cloud-occurrence}. The analysis
shows that the total cloud fraction (determined as the fraction of profiles
with clouds at any height in the lidar cloud mask) is underestimated in ICON by
about 10\% and in the reanalyses by about 20\%. When analysed by height, ICON
overestimates cloud occurrence below 1 km and underestimates it above, MERRA-2
underestimates cloud occurrence at all heights by up to about 10\%, especially
near the surface, and ERA5 simulates cloud occurrence relatively well above 1
km, but strongly underestimates it near the surface.  We note that fog or
near-surface clouds are strongly underestimated in the reanalyses (fog and
clouds are both included in the cloud occurrence).  As shown in Fig.
\ref{fig:cloud-occurrence-panel}, the biases are relatively consistent across
the data sources and longitudes. We conclude that the ICON results  the
observations better than the reanalyses in this metric.

For all observations considered (Fig. \ref{fig:cloud-occurrence}a), the data
show cloud occurrence peaking at the surface, whereas the models show a higher
peak (at about 500 m). The models generally underestimate the total cloud
fraction by 10--20\% and show a strong reduction in cloud occurrence near the
surface, which is not supported by the observations. ICON and ERA5 overestimate
cloud occurrence at their peak (between 0--1 km). Above 1 km, ICON and MERRA-2
underestimate cloud occurrence, but ERA5 is accurate to about 3\% or less. The
exaggerated peak in models is partly supported by the lifting condensation
level (LCL) distribution, which peaks higher by about 200 m in the models than
in the observations (at the surface), although this is not very pronounced.
This is indicative of near-surface relatively humidity being often close to
saturation in the observations but not in the models.

@@

When subsetted by cyclonic and non-cyclonic situations (Fig.
\ref{fig:cloud-occurrence}d, e), we see that the cyclonic situations have a
larger amount of observed cloudiness, including peak and total cloud fraction,
both by about 7\%.  In the cyclonic situations, the model vertical profiles of
cloud occurrence compare well with observations, but they peak higher by about
200 m and stronger by about 8\%. The reanalyses still tend to underestimate
cloud occurrence above 1 km by about 5\% and near the surface by about 14\%.
Non-cyclonic situations are similar to the general case, partially also because
they form the majority of cases.

When subsetted by conditions of weak and strong stability (Fig.
\ref{fig:cloud-occurrence}f, g), as defined in Section
\ref{sec:cyclone-stability}, we see that in situations of strong stability
cloud occurrence peaks strongly at the surface in observations, compared to
situations of weak stability, where the peak is more diffuse between 0 and 1
km.  It is work mentioning that conditions of strong stability might be
associated with the formation of advection fog, such as in situation of warm
air advection from the north over colder sea surface, thus inducing fog
formation by cooling of the warm and humid air by the cold surface.  In
situations of strong stability, the models have smaller biases than in weak
stability, with overestimated peak by up to 12\%, underestimated cloud
occurrence above 1 km by up to 5\% and underestimated cloud occurrence near the
surface by about 11\%.  In situations of weak stability, the bias in ICON is
very pronounced, with a much stronger peak at about 500 m, ERA5 underestimates
cloud occurrence below 1 km (especially near the surface), and MERRA-2
underestimates cloud occurrence even more strongly.

@@

In Fig.  \ref{fig:cloud-cover} we show the results for the same subsets of data
as in Section \ref{sec:cloud-occurrence}. Observations display the highest
proportion of high cloud cover values (5--8 oktas), peaking at 7 oktas. This
pattern is not represented by ICON or either reanalyses.  While ICON is closest
to matching the observed distribution, it tends to be 1 okta clearer than the
observations, peaking at 6 oktas, and substantially underestimating days with 8
oktas.  Overall, the reanalyses show results similar to each other,
underestimating cloud cover by about 2 oktas and strongly underestimating days
with 7 and 8 oktas. Of the two reanalyses, MERRA-2 has slightly higher cloud
cover then ERA5, by about 6\% at 6 octas, which makes it more consistent with
observations.

@@

Overall, the mean $\theta_v$ is accurate less than 0.5 K in ICON and MERRA-2,
being only colder by up to 2 K in the mid-to-high troposphere (less stable) in
ICON than in observations (Fig. \ref{fig:potential-temperature}a).  Larger
differences exist, however, in the 40--55°S zone, where ICON is colder by about
5 K at higher altitudes (Fig.  \ref{fig:potential-temperature}b). In other
subsets, the bias is relatively small. MERRA-2 is very close to the
observations, possibly due to a high accuracy of assimilation of this quantity.
Notably, the variability of virtual potential temperature (as represented by
the percentiles) is much smaller in ICON than in the observations. This
indicates that the model's internal variability in the lower-tropospheric
thermodynamic conditions in the SO is smaller than in reality.
